---
title: "Student_Perfomance"
author: "Purity"
date: "29/05/2020"
output: github_document
    
---
**Given below are attributes**

*Attributes for both student-mat.csv (Math course) and student-por.csv (Portuguese language course) datasets:*

1 school - student's school (binary: 'GP' - Gabriel Pereira or 'MS' - Mousinho da Silveira)

2 sex - student's sex (binary: 'F' - female or 'M' - male)

3.age - student's age (numeric: from 15 to 22)

4.address - student's home address type (binary: 'U' - urban or 'R' - rural)

5.famsize - family size (binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3)

6.Pstatus - parent's cohabitation status (binary: 'T' - living together or 'A' - apart)

7.Medu - mother's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)

8.Fedu - father's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)

9.Mjob - mother's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')

10.Fjob - father's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')

11.reason - reason to choose this school (nominal: close to 'home', school 'reputation', 'course' preference or 'other')

12.guardian - student's guardian (nominal: 'mother', 'father' or 'other')

13.traveltime - home to school travel time (numeric: 1 - 1 hour)

14.studytime - weekly study time (numeric: 1 - 10 hours)

15.failures - number of past class failures (numeric: n if 1<=n<3, else 4)

16.schoolsup - extra educational support (binary: yes or no)

17.famsup - family educational support (binary: yes or no)

18.paid - extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)

19.activities - extra-curricular activities (binary: yes or no)

20.nursery - attended nursery school (binary: yes or no)

21.higher - wants to take higher education (binary: yes or no)

22.internet - Internet access at home (binary: yes or no)

23.romantic - with a romantic relationship (binary: yes or no)

24.famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent)

25.freetime - free time after school (numeric: from 1 - very low to 5 - very high)

26.goout - going out with friends (numeric: from 1 - very low to 5 - very high)

27.Dalc - workday alcohol consumption (numeric: from 1 - very low to 5 - very high)

28.Walc - weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)

29.health - current health status (numeric: from 1 - very bad to 5 - very good)

30.absences - number of school absences (numeric: from 0 to 93)
These grades are related with the course subject, Math or Portuguese:
1.G1 - first period grade (numeric: from 0 to 20)

2.G2 - second period grade (numeric: from 0 to 20)

3.G3 - final grade (numeric: from 0 to 20, output target)

Additional note: there are several (382) students that belong to both datasets .

These students can be identified by searching for identical attributes

that characterize each student, as shown in the annexed R file.
Source Information

P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7.
Fabio Pagnotta, Hossain Mohammad Amran.
Email:fabio.pagnotta@studenti.unicam.it, mohammadamra.hossain '@' studenti.unicam.it
University Of Camerino

<https://archive.ics.uci.edu/ml/datasets/STUDENT+ALCOHOL+CONSUMPTION>


# ***Business Understanding***
A Non Governmental Organization has collected data from students doing math and portugese course in high school.They would like to gain insights from the social,gender and study information about them.
They would also like you to predict the student final results.

* Solution

Do EDA and get as many insights as possible.Use the suitable supervised machine learning algorithm to predict what is needed.

# ***Data Understanding***
Here we load the datasets needed.Know about it all by checking distributions/skewness,shape and skewness.
```{r}
#load the data sets and libraries
library(tidyverse)##Is a set of packages that include dplyr,tidyr etc
d1=read.csv("student-mat.csv",header=TRUE)
d2=read.csv("student-por.csv",header=TRUE)
head(d2)
d4=merge(d1,d2,by=c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob","reason","nursery","internet"))
#d4[!duplicated(d4$guardian.x,d4)]
print(nrow(d4))


```
AS seen above the two are merged to give a common dataset with 382 rows
```{r}
d3<-read.csv("d3.csv")
view(d3)
```


## *Check The Shape*
```{r}
dim(d3)
anyDuplicated(colnames(d3))

```
## *Check the structure*
```{r}
#I normally use glimpse 
glimpse(d3)
```
## *next check for missing values*
```{r}
#Check per each column
colSums(is.na(d3))
```


## *Have a summary of all*
```{r}
summary(d3)
```

## *Remove Some things*
since this is a merge from portugese and math course for students who do both they have final grade for the two 
So we have to have common final grade
I decided to add and divide by 2 to get the common final grade for a student
```{r}
d3<-d3%>%
  mutate(G1=(G1.x +G1.y)/2,
         G2 = ( G2.x  + G2.y  )/2,
         G3 = ( G3.x + G3.y  )/2)
```
know i will drop the grades from both.
```{r}
dim(d3)
```
```{r}
d3<-select(d3,-c(30:35))
```

```{r}
str(d3)
```

```{r}
d3[27:32]
```


## *now lets set text to be uniform*
```{r}
colnames(d3)<-gsub(".x","",colnames(d3))
```
```{r}
```

```{r}
ncol(d3)
colnames(d3)[2]<-"sex"
```

## *now check place them to same case*
```{r}
colnames(d3)<-tolower(colnames(d3))
```

```{r}
library(GGally)
colnames(d3)
numeral<-d3%>%
  select_if(is.factor)
glimpse(numeral)

  #cor()%>%
  #corrplot::corrplot()
```
# ***EDA***
## **Univariate**
### *numerics*
```{r}
#par(mfrow=c(2,2))
numeral_n<-d3%>%
  select_if(is.numeric)
glimpse(numeral_n)

```
```{r}
for(i in 1:16) {
hist(numeral_n[,i], main=names(numeral_n)[i],xlab =names(numeral_n)[i],col = "powder blue" )
}
```

***From the above there are lots of conclusion you can make about high school students.***

**age ranges from 15-20 though there is these student with 22**

**Most parents for these kids are educated** 

**For absentinsm there seem to a student who has been absent for 60 days do abnormal.**
**As you can also see these students some of them take alcoholo weekdays but more on weekends.**

### *For factors*
```{r}

for(i in 1:16) {
counts <- table(numeral[,i])
name <- names(numeral)[i]
barplot(counts, main=name,col = "red")

}
```

## **Multivariate Analysis**

**Check for multicollinearlity**
```{r}
x=numeral_n[,1:15]
y=numeral_n[,16]

numeral_n%>%
  gather(-g3,key = "var",value = "value")%>%
  ggplot(aes(x = value,y =g3 ))+
  facet_wrap(~var,scales = "free")+
  geom_jitter()+
  geom_smooth(method = "lm")
```

***I realised the data is given in a nominal form ***

**Even so i didn't change them factors**

**I also realised the grades for g1,g2,g3 are very strongly correlated**

**absentism can obviuosy be a reason for failure**

**Also the more the failing the lower the grade**

**Alcohol consumption can greatly influence the final grade**


```{r}

# This sets plot images to a nice size.
options(repr.plot.width = 4, repr.plot.height = 4)
ggplot(d3,aes(x=g3,y =absences))+
  geom_col(fill = "blue")+
  facet_wrap(~internet+address)+
  theme_classic()


```
***From the plot above:**

**Most students perfomance is influenced with whether the internet is available**

**Most students live in the urban settlement these also influences there perfomance**
```{r}
options(repr.plot.width = 4, repr.plot.height = 4)
ggplot(d3,aes(x=g3,y =failures))+
  geom_col(fill = "blue")+
  facet_wrap(~higher)+
  theme_classic()

```

**Those perfoming lower have no ambitions of going to higher learning**
```{r}
library(caret)
set.seed(107)
inTrain <- createDataPartition(
  y = d3$g3,
  ## the outcome data are needed
  p = .75,
  ## The percentage of data in the
  ## training set
  list = FALSE
)
training <- d3[ inTrain,]
testing  <- d3[-inTrain,]

```


# **Modelling**

## *Prepare the data for modelling/preprocessing*
### *Dummyfy the data*

```{r}
training<-model.matrix(~.,data = training)
```
```{r}

testing  <- model.matrix(~.,data = testing)
```

### *Scale the data*

```{r}
library(caret)
library(gbm)
# load data
d3_t<-preProcess(training,method = "YeoJohnson")
training<-as.data.frame(predict(d3_t,newdata = training))
head(training)
```


## *split the data into train and test then train*
```{r}
trainControl <- trainControl(method="repeatedcv", number=5)
set.seed(7)
#usage of caret package and caret available model
fit <- train(g3~g1+g2+absences+failures+studytime+famrel+dalc+activitiesyes+	traveltime+famsupyes, data=training, method="gbm", metric="RMSE", trControl=trainControl)

```

### *check the tuning*

```{r}
fit$finalModel$tuneValue
fit$modelInfo$parameters
fit$results
```
```{r}

fit$bestTune
```

## *Predict the future data/testing data*
```{r}
testing$pred<-predict(fit,testing)
head(testing$pred)

```

### *check for variample importance*
```{r}
library(caret)
summary(fit)
```

## *Conclusion*

1.The second grade is the most important factor that would determine a student's final.
2.Absentism is a factor also the students who are mostly absent are the likely ones to fail in their final grade.
3.A factor like how students use their time also can be a concern in perfomance.

### *Recommendations*
Students should work on every exam and perfom well so that it can influence their final grade.Teachers should know what makes their students be absent and help them.
They should also advise them on time management.Also parents should be concerned on their children perfomance.For example they should organize good transport so that the travel time should be little.

